---
title: "Lab Name"
author: "Author Name"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

# Load the packages
- Check the list of the required packages
- change the path as appropriate

```{r load-packages, message=FALSE}
rm(list=ls())
library(openintro)

knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = normalizePath("C:\\Users\\kathm.alismail\\Downloads"))
#setwd("C:\\Users\\kalis\\Downloads\\Harrisburg University\\502\\Project") 
required.packages <-c("tidyverse","dplyr","ggplot2","ggpubr","knitr","devtools","kableExtra","data.table","ggsci","gridExtra","RColorBrewer", "ggthemes","ggsci","rio","reshape")
#"lmerTest","dplyr"
#ls("package:ggplot2") #list functions of a package
# Load required packages
lapply(required.packages, library, character.only = TRUE)

```

### Load Data

- load data by either (a) "Import Dataset" (b) read.csv function

```{r Load & Analyze}
# Load and view Data
Data<-read.csv("Exce Name.csv",header=TRUE, fileEncoding="UTF-8-BOM") 
attributes(Data)
summary(Data)
str(Data)                 # structure 
table(Data$level)         #for catagorical & Variables  
ls(Data)
class(Data)
unique(Data)              # gives levels
```

### Creating a summary Table
_ change the variable names and the categorical factors

...
```{r Summary Tabel}
#Create summary tables
     New_Name <- Dataset_table_Name %>%                           # Get #mean & standard deviation by group
    group_by("Catagorical Colmn Name") %>%
    summarise_at(vars("Variable Column Name"),list(mean = mean, sd = sd)) %>% 
    as.data.frame()

  New_Name
  
  
  # Another grouping 
  Agg=aggregate("Variable Column Name", list("Catagorical Colmn Name"), FUN=sd,mean) # Grouping using aggregate function


```

### Graphing 
1. Make sure there is a catagorical column

Set ggplot color manually:

scale_fill_manual() for box plot, bar plot, violin plot, dot plot, etc
scale_color_manual() or scale_colour_manual() for lines and points
```{r Line Graph}
Var1= 1:30
Var2= seq(12,70,by=2)
CategoricalVar = rep(LETTERS[1:3], each = 10)
groups=rep(c("One","Two"), each = 15)
Data = data.frame(Var1, Var2, CategoricalVar,groups)
Data$groups <- as.factor(ifelse(Data$Var1<7, 'One',
                     ifelse(Data$Var1<15, 'Two',
                     ifelse(Data$Var1<24, 'Three',"Four"
                     ))))

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442","#FFDB6D", "#0072B2", "#D55E00", "#CC79A7")
```
Line Plots with Groups
geom_line(aes(y = Var2, group = CategoricalVar, color = CategoricalVar), size = 1)
 scale_color_nejm()+
 scale_color_igv()
```{r Setting up ggplot}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
theme_set(
  theme_minimal() +
    theme_bw()+
    theme(plot.title = element_text(face = "bold",hjust = 0.5),
          axis.title=element_text(face="bold"),
          plot.subtitle = element_text(hjust = 0.5),
          legend.title = element_text(face="bold"),
          axis.title.y = element_text(margin = margin(t = 0, r = 10, b = 0, l = 0)),
          axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0))) #theme for the title (bold and middle), axis , and axis position (margin)
  )
```
 
 
 

```{r}
## Normal GGplot Line with groupds
  ggplot(Data, aes(x = Var1, y = Var2)) +
    geom_line(aes(group = CategoricalVar, color = CategoricalVar), linewidth = 1)+
  stat_cor(aes(label = paste(..rr.label.., ..p.label.., sep = "~`,`~"),color = CategoricalVar), label.x = 4)+
  facet_wrap(~CategoricalVar)+
    labs(x = "X-Axis", y = "Y-Axis", title = "Title", face = "bold",tag = "Figure 1")+
    #scale_color_jco()+
    scale_color_manual("Legend",values = cbPalette ,labels = c("A","B","C"))
    
```


## Scatterplots
```{r}
  ggplot(Data, aes(x = Var1, y = Var2)) +
    geom_point(aes(group = CategoricalVar, fill = CategoricalVar), size = 3, pch=21, alpha=0.8, color="black")+
    labs(x = "X-Axis", y = "Y-Axis", title = "Title", face = "bold", tag = "Figure 2")+
    #scale_color_igv()+
    scale_fill_manual("Legend",values = cbPalette ,labels = c("A","B","C")) # *This always need pch*
  #stat_smooth(method = lm, se = FALSE,color="black")
    #scale_color_brewer(palette="jco")
    scale_fill_jco()
```



```{r Graphing Box plot}
#Plotting
ggplot(Data, aes(x=CategoricalVar, y=Var1, fill=CategoricalVar))+
  geom_boxplot()+
  labs(x = "X-Axis", y = "Y-Axis", title = "Title", face = "bold" ,tag = "Figure 3")+
  scale_fill_manual("Legend",values = cbPalette ,labels = c("A","B","C"))
```


```{r Bar plot}
ggplot(Data,aes(x=CategoricalVar,y=Var1,fill=CategoricalVar))+
    stat_summary(fun = "mean", geom = "bar", position = "dodge")+
    stat_summary(fun.data = "mean_sdl", geom = "errorbar", position = position_dodge(0.9), width= 0.3) + 
  scale_fill_manual("Legend",values = cbPalette ,labels = c("A","B","C"))


ggplot(Data,aes(x=CategoricalVar,y=Var1,fill=groups))+
    stat_summary(fun = "mean", geom = "bar", position = "dodge")+
    stat_summary(fun.data  = "mean_sd", geom = "errorbar", position = position_dodge(0.9), width= 0.3) + 
  scale_fill_manual(name="GROUPS",values = cbPalette)
  scale_fill_jco()



#Creating summary Data
New_Data <- Data %>%                           # Get #mean & standard deviation by group
    group_by(CategoricalVar,groups) %>%
    summarise_at(vars(Var1),list(mean = mean, sd = sd)) %>% 
    as.data.frame()


  # Another grouping 
  New_Data1=aggregate(Var1, list(CategoricalVar), FUN=mean) # Grouping using aggregate function
  


ggplot(New_Data, aes(x=CategoricalVar, y=mean, fill=groups))+
  geom_bar(stat="identity",colour="black",position = "dodge")+
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd),position = position_dodge(0.9),width= 0.3)




```

## Important R functions
- Computation & Automtion 
- apply() This a function that applies any function on different colmns/rows 
```{r apply}
B=apply(Data[,1:2],2,mean)
B[1]
A=lapply(Data,mean)
A$Var1
C=sapply(Data[,1:2],mean)
C[1]
D=tapply(Data$Var1,list(Data$CategoricalVar),mean) #***
D[1]
```

## More on Automations

# for loop

```{r for loop}
for(i in 1:length(Data$Var1)){
  Data$Var3[i] <- Data$Var1[i]^2
}
```

```{r Correlations}
#Correlations
Data_sub=Data[,-c(3,4)]
install.packages("ggcorrplot")
library(ggcorrplot)
corr <- round(cor(Data_sub), 1)
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE)

data(mtcars)
corr <- round(cor(mtcars), 1)
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE)
mpg
```
## Hypothesis Testing

#inference 
```{r}
#group by 
by(Data$Var1, Data$CategoricalVar, mean)

inference(y = nc$weeks, x = nc$marital, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical") 

inference(y = Data$Var1, x = Data$CategoricalVar, est = "mean", type = "ht", null = 0, alternative = "greater", method = "theoretical") 
# type = "ht" hypothesis, "ci" confidence interval


```


# Linearity 
```{r}
lm1 <- lm(Var2 ~ ., #Y is predicted by all variables in the data
           data = Data) #You can use categorical variables now!
summary(lm1)

lm2 <-lm(Var2 ~ Var1, data=Data)
model.diag.metrics=augment(lm1)
ggplot(model.diag.metrics, aes(Var1, Var2)) +
  geom_point() +
  stat_smooth(method = lm, se = FALSE) +
  geom_segment(aes(xend = Var1, yend = .fitted), color = "red", size = 0.3)

lm1_standardized <- rstudent(lm1) #Standarized (residuals/sd)
summary(lm1_standardized)

fitvalues <- scale(lm1$fitted.values) #standarized fitted value (x-mean/sd) where x is predicted value


{qqnorm(lm1_standardized)
abline(0,1)}

{qqnorm(residuals(lm1))
}
```
# Normality 
```{r}
library(moments)
skewness(Data[ , -c(3,4)])
#A positive skewness value indicates a right-skewed distribution (tail extends to the right).
#A negative skewness value indicates a left-skewed distribution (tail extends to the left).
#A skewness value close to zero suggests a relatively symmetric distribution.

kurtosis(Data[ , -c(3,4)]) - 3 #to get excess kurtosis

#Normality of the model 
hist(residuals(lm1))
hist(lm1_standardized, breaks=15)

#shapiro Walks Normality Test 
data_groups <- split(Data$Var1, Data$CategoricalVar)
# Perform Shapiro-Wilk test for each group
shapiro_results <- lapply(data_groups, shapiro.test)
sapply(shapiro_results, function(x) x$p.value)

# Homoge
plot(lm1$fitted.values,residuals(lm1))
abline(h=0)

# standarized residual plot "
{plot(fitvalues, lm1_standardized) #standarized predicted value and standarized residuals
abline(0,0)
abline(v = 0)}
```


## Find the best fit Model 


```{r}
step(lm1, direction = "backward")

```

```{r Power in t test}
library(pwr)

t.test(PAL_cell ~ gender, data = Data, 
       var.equal = FALSE, paired = FALSE)
pwr.t.test(n = NULL, #leave NULL
           d = effect$d, #effect size
           sig.level = .05, #alpha
           power = .80, #power 
           type = "two.sample", #independent
           alternative = "two.sided") #two tailed test
```


```{r}
#dplyer
Summary_Data <- Data %>%
  filter(groups=="One") %>%
  select(Var1,Var2,CategoricalVar) %>%
  mutate(Var3 = Var1 + Var2) %>%
  arrange(desc(Var3))
  
Summary_Data2 <- Data %>%
  group_by(CategoricalVar) %>%
  summarise(avg1=mean(Var1),avg2=mean(Var2))

Var3= 1:30
Var4= seq(12,70,by=2)
CategoricalVar2 = rep(LETTERS[1:3], each = 10)
groups=rep(c("One","Two"), each = 15)
Data2 = data.frame(Var3, Var4, CategoricalVar2,groups2)

Summary_Data %>%
  inner_join(Data2, by = c("groups" = "group"))
```

